Hunting experience shapes individual foraging specialisation and predator-prey interactions in an online videogame

### 2023-02-17 manuscript/Chapter2-manuscript.Rmd
started writting the discussion. 1st parargraph

@@

The development of foraging tactics is crucial for young predators to reach adulthood and survive [@Phillips.etal2017;@Heithaus.etal2018]. Theory predicts that foraging specialization may emerge via learning, limitations in memorizing multiple complex hunting skills, and expertise [@Tinker.etal2009;@Dukas2019]. @Dukas2019 defines expertise as the characteristics, skills, and knowledge allowing individuals with extensive experience to outperform novices on complex tasks. This body of work suggests that the development of expertise is an optimizing process that promotes foraging specialization. Empirical studies on human and non-human hunters show that experience optimizes the efficiency (e.g. search and handling times, return rates) of their foraging tactics potentially via associative images or reliance on prey and environmental cues [@Edwards.Jackson1994;@Morse2000;@macdonaldCrossculturalComparisonLearning2007;@Reid.etal2010;@Wilson-Rankin2015]. Such optimization may thus reinforce the use of the same tactic (i.e. specialization) if its success is constant each time a prey is encountered. It may also be costly to attempt different hunting tactics by trial and error when prey are scarce or highly unpredictable [@Dukas1998;@Estes.etal2003;@Mery.Burns2010]. An alternative mechanism is that the costs/risks faced by hunters when switching foraging tactics may be offset by gaining experience and information on their prey, leading to increased individual foraging flexibility [@Ishii.Shimada2010;@Mery.Burns2010;Kelley.Magurran2011;@Snell-Rood2013]. To develop proper responses to fluctuating resources, individuals would need to sample their environment broadly during an extensive period of time, followed by higher performance at later stages of development [reviewed in @Snell-Rood2013]. However, individuals may face different challenges during their lifetime, which could lead to both mechanisms operating at the same time within a population.

@@

We used the default number of basis functions (K) in "brms" for the models to estimate the relationship between hunting success and experience. We assumed that the random intercepts for the predator ID ($id$) followed a Gaussian distribution with estimated standard deviation ($id \sim N(0,\ I\ \sigma_{id})$). We estimated the parameters in all models using four MCMC chains with 1000 posterior samples for each parameter. To do so, we ran 2500 iterations for the first model with a thinning set to eight, and 1500 iterations with a thinning set to four for the second and third models; burn-in was set to 500 iterations in each model. We provide a description of how we specified the priors for each model in the supporting information. After fitting the three models, we proceeded to select the one with the best predictive accuracy using approximate leave-one-out cross-validation with Pareto-smoothed importance sampling [@Vehtari.etal2017;@Piironen.Vehtari2017;@Vehtari.etal2022].

@@

Our results confirm that individual predators displayed important differences in the development of their expertise, as the best model was the one with group-level smoothers exclusively (Table I). For instance, with experience, some predators steadily increased their hunting success while others displayed a decrease in prey capture (Figure 1A-S1). Otherwise, there were players that maintained either a high or low success across their experience (Figure 1A). Another pattern that emerged was that some individuals optimized their success at an experience level that was below what we considered as "advanced" (i.e. >300 matches played).

@@

Outline of the main results

## Individuals differ in the development of their expertise

## Specialization does not change with experience at the population level

## Individuals either specialize or increase their flexiblity with experience

## Experience strenghtens predator-prey interactions

## Experience dampens the relationship between behaviour and success

>>>>

The development of foraging tactics is crucial for young predators to reach adulthood and survive [@Phillips.etal2017;@Heithaus.etal2018]. Theory predicts that foraging specialization may emerge via learning, limitations in memorizing multiple complex hunting skills, and expertise [@Tinker.etal2009;@Dukas2019]. @Dukas2019 defines expertise as the characteristics, skills, and knowledge allowing individuals with extensive experience to outperform novices on complex tasks. This body of work suggests that the development of expertise through extensive practice is an optimizing process that promotes foraging specialization. Empirical studies on human and non-human hunters show that experience optimizes the efficiency (e.g. search and handling times, return rates) of their foraging tactics potentially via associative images or reliance on prey and environmental cues [@Edwards.Jackson1994;@Morse2000;@macdonaldCrossculturalComparisonLearning2007;@Reid.etal2010;@Wilson-Rankin2015]. Such optimization may thus reinforce the use of the same tactic (i.e. specialization) if its success is constant each time a prey is encountered. It may also be costly to attempt different hunting tactics by trial and error when prey are scarce or highly unpredictable [@Dukas1998;@Estes.etal2003;@Mery.Burns2010]. An alternative mechanism is that the costs/risks faced by hunters when switching foraging tactics may be offset by gaining experience and information on their prey, leading to increased individual foraging flexibility [@Ishii.Shimada2010;@Mery.Burns2010;Kelley.Magurran2011;@Snell-Rood2013]. To develop proper responses to fluctuating resources, individuals would need to sample their environment broadly during an extensive period of time, followed by higher performance at later stages of development [reviewed in @Snell-Rood2013]. However, individuals may differ in the challenges that they face during their lifetime, which could lead to both mechanisms operating at the same time within a predator population.

@@

@@

We used the default number of basis functions (K) in "brms" for the models to estimate the relationship between hunting success and experience. We assumed that the random intercepts for the predator ID ($id$) followed a Gaussian distribution with estimated standard deviation ($id \sim N(0,\ I\ \sigma_{id})$). We used Gaussian priors for the game duration ($N(0,\ 1)$), the intercept ($N(0,\ 0.5)$), and the precision paremeter ($N(2,\ 0.5)$). We employed the default Student t priors for the smoothing parameter ($t(3,\ 0,\ 2.5)$). In all models, we specified four MCMC chains to estimate the parameters, with 1000 posterior samples for each parameter. To do so, we ran 2500 iterations for the first model with a thinning set to eight, and 1500 iterations with a thinning set to four for the second and third models; burn-in was set to 500 iterations in each model. After fitting the three models, we proceeded to select the one with the best predictive accuracy using approximate leave-one-out cross-validation with Pareto-smoothed importance sampling [@Vehtari.etal2017;@Piironen.Vehtari2017;@Vehtari.etal2022].

@@

Our results confirm that individual predators displayed substantial differences in the development of their expertise, as the best model was the one with group-level smoothers exclusively (Table I). For instance, with experience, some predators steadily increased their hunting success while others displayed a decrease in prey capture (Figure 1A-S1). Otherwise, there were players that maintained either a high or low success across their experience (Figure 1A). Another pattern that emerged was that some individuals optimized their success at an experience level that was below what we considered as "advanced" (i.e. >300 matches played).

@@

For decades, hunting experience has been hypothesized to play a major role in predator-prey interactions []. Yet, the lack of empirical evidence has led the existing literature to predict conflicting outcomes []. It has thus remained unclear whether predators specialize in their tactic with experience, how such specialization emerges depending on the type of prey that predators encounter, and whether specialization with experience leads to an increase in hunting success. We adressed these questions by capitalizing on an online videogame system simulating a predator-prey interaction, where both the predator and the prey behaviours are simultaneously monitored. Our results show that experience does not generate responses at the population level. Instead, experience appears to be a property of individuals, such that each predator developed, through extensive practice, their own behavioural trajectory. Individuals became increasingly distinct hunters with experience; some specialized at fast-paced hunting, while others adopted a slower and more flexible tactic. The predators' tactics were also associated with prey encounters, such that those that adopted a specialized cursorial tactic experienced more predictable encounters, while slower and flexible hunters experienced more unpredictable encounters, with no difference in success between the two types of hunters. With experience, this predator-prey behavioural relationship grew stronger, while the relationship between the prey's behaviour and hunting success diminished. Together, our observations suggest that predators use their experience to adjust their hunting tactic, and the degree at which they specialize in it, based on the prey that they encountered throughout their lifetime. Moreover, the stronger predator-prey interaction through experience may imply the emergence of an arms race mediated through learning.

## The role of hunting experience on the predator's foraging behaviour

- Individuals differ in the development of their expertise

- Specialization does not change with experience at the population level

- Individuals either specialize or increase their flexiblity with experience

*This has important implications for future theoretical models predicting a directional increase in specialization or flexibility. should incorporate individual variation if both mechanisms operate at the same time*

## Predator-prey interactions change with experience

- Experience strenghtens predator-prey interactions

- Experience dampens the relationship between behaviour and success
